Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-02-13 ribose-paper.RMD
updates to introduction

Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress. Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired growth (ALVAREZ18;MANN14;THALACKER13). Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu? -->
Although RT program modifications such as exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020]<!-- Add ref for more rt program modifications -->
changes in e.g., training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020]. Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory? -->

@@

Ribosome biogenesis is suggested to be in part regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18). Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09). This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19;TANAKA18). Upon stimulation by mTORC1, assembly of the preinitiation complex (PIC) marks the initiation of ribosome biogenesis (VONWALDEN19;FIGUEIREDO19). The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19). Further, the upstream binding factor (UBF) and c-Myc have vital roles in the initiation, where UBF binds the PIC to the rDNA promoter and c-Myc as a transcription factor (FIGUEIREDO19, VONWALDEN19). Pol I transcribes the 47S pre-rRNA, suggested as the rate-limiting step in ribosome biogenesis (VONWALDEN19, FIGUEIREDO19). The 47S pre-rRNA is processed into the 18S, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19), while the last rRNA 5S is transcribed by Pol III, and the ribosomal proteins by Pol II (MULLINEUX12, VONWALDEN19). Importantly, these events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, suggested to predict long term adaptations to RT (HAMMARSTROM21). Since a blunted ribosome biogenesis has been observed to reduce/inhibit RT responses, despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and/or nutritional signals may ameliorate this. Previously, high glucose treatment promoted ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation in mice and cell cultures (MARIAPPAN11;TANAKA15;ZHAI12), making glucose an interesting player in this perspective.

Thus, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (TANAKA18), it is unclear whether there are acute effects on ribosome biogenesis *per se* with ingesting glucose before and after RT. Further, there are equivocal observations on insulin's impact on ribosome biogenesis and MP, where previous meta-analyses suggest insulin to act in concert with amino acid intake to enhance MPS, and insulin may reduce skeletal muscle protein breakdown when amino acids are scarce (ABDULLA16). Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks. They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session. Our main purpose was to investigate whether there are acute effects of glucose- compared to placebo ingestion before and after RT. To assess this purpose, we measured changes in ribosome biogenesis as accumulation of total RNA, expression of the four mature rRNA's (18S, 5.8S, 28S, 5S), 47S pre-rRNA, and levels of proteins associated with transcription initiation (c-Myc, UBF, rpS6?). As secondary outcomes, we measured the development in skeletal muscle performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque. We hypothesised that RT would indeed lead to a prominent accumulation of total RNA, increased expression of rRNA and associated proteins. Further, we hypothesised that RT with glucose and protein would induce an enhanced response compared to placebo, in markers of ribosome biogenesis, as well as muscular performance and -recovery.

>>>>

Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress. Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired muscle growth (ALVAREZ18;MANN14;THALACKER13). Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu? -->
Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020].<!-- Add ref for more rt program modifications -->
For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020]. Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory? -->

@@

Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995]. Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18). mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels (HOPPE09). Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000]. mTOR contributes to forming the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19). Parallel to mTORC1 [@Mori_2021; West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011]. However, interestingly, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011]. Additionally, independently from UBF, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription[@Zhai_2012]. Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.

The purpose of this investigation was to determine the effect of 
<!-- -->

the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks. They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session. Our main purpose was to investigate whether there are acute effects of glucose- compared to placebo ingestion before and after RT. To assess this purpose, we measured changes in ribosome biogenesis as accumulation of total RNA, expression of the four mature rRNA's (18S, 5.8S, 28S, 5S), 47S pre-rRNA, and levels of proteins associated with transcription initiation (c-Myc, UBF, rpS6?). As secondary outcomes, we measured the development in skeletal muscle performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque. We hypothesised that RT would indeed lead to a prominent accumulation of total RNA, increased expression of rRNA and associated proteins. Further, we hypothesised that RT with glucose and protein would induce an enhanced response compared to placebo, in markers of ribosome biogenesis, as well as muscular performance and -recovery.
